---
title: 'HW_07'
author: "Cody Grainger"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: inline
---

```{r}
library(tidyverse)
```

```{r}
lim_levels <- c("PBL","KNRM","VBG","WP","PF","DP","SIO","CTZ1","CTZ2","CNM" )

limpetz <- read_csv("data/limpets.csv", col_types= cols(
  Length = col_number(),
  Site = col_character(),
  Protected = col_character()
))
limpetz

ggplot(data = Limpets, mapping = aes(x = Length, y = Site, color= Protected)) + 
  geom_boxplot(xlab="Size(mm)",ylab="Site")
```
The Data is already tidy for use. According to the plot, sites 'WP' and 'DP' have outliars. The data appers to support limpets in protected zones having greater lengths/sizes than those in unprotected zones. 

```{r}
data("roseate_terns")
roseate_terns
terns <- read_tsv("data/roseate_terns.txt")

terns2 <- drop_na(terns,AR,AL,AC,BR,BL,BC,JR,JL,JC)
terns2

terns3 <- terns2 %>%
  pivot_longer(c('AR','AL','AC','BR','BL','BC','JR','JL','JC'),names_to="Site",values_to="Population")
terns3

terns3 %>%
  ggplot(mapping=aes(x=year,y=Population,color=Site,position="jitter"))+
  geom_point()

terns3 %>%
  ggplot(mapping=aes(x=year,y=Population,color=Site))+
  geom_point()+
  facet_wrap(~Site,ncol=3)
```
Populations on site AR, AL, and BL appear to show the most significant increases with AR being the largest and most dramatic increase. JR seems to have had a population dcline. The noncontinuous plots have had some data rows dropped if there were any NA values within the row.




