---
title: 'HW_07'
author: "Cody Grainger"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: inline
---

```{r}
library(tidyverse)
```

## Limpets
```{r}
#Define levels
lim_levels <- c("PBL","KNRM","VBG","WP","PF","DP","SIO","CTZ1","CTZ2","CNM" )


limpetz <- read_csv("data/limpets.csv", col_types= cols(
  Length = col_number(),
  Site = col_character(),
  Protected = col_character()
))
limpetz

#Boxplot of limpet data
ggplot(data = Limpets, mapping = aes(x = Length, y = Site, color= Protected)) + 
  geom_boxplot(xlab="Size(mm)",ylab="Site")
```
The Data is already tidy for use. According to the plot, sites 'WP' and 'DP' have outliars. The data appers to support limpets in protected zones having greater lengths/sizes than those in unprotected zones. 

## Roseate Terns
```{r}
# Load data
data("roseate_terns")
roseate_terns
terns <- read_tsv("data/roseate_terns.txt")

terns2 <- drop_na(terns,AR,AL,AC,BR,BL,BC,JR,JL,JC)
terns2

terns3 <- terns2 %>%
  pivot_longer(c('AR','AL','AC','BR','BL','BC','JR','JL','JC'),names_to="Site",values_to="Population")
terns3

# Scatterplot
terns3 %>%
  ggplot(mapping=aes(x=year,y=Population,color=Site,position="jitter"))+
  geom_point()

terns3 %>%
  ggplot(mapping=aes(x=year,y=Population,color=Site))+
  geom_point()+
  facet_wrap(~Site,ncol=3)
```
Populations on site AR, AL, and BL appear to show the most significant increases with AR being the largest and most dramatic increase. JR seems to have had a population dcline. The noncontinuous plots have had some data rows dropped if there were any NA values within the row.

## Blacklip Abalone
```{r}
#Setting up read.csv
regular_ol_babalone <- read_csv("data/abalone.csv")
regular_ol_babalone
abalone2 <- select(regular_ol_abalone,- X1)

#Test/Experiment
abaloneee <- read_csv("data/abalone.csv",skip=1,col_names=TRUE)
abaloneee

#The boxplot
ggplot(data=abalone2)+
  geom_boxplot(mapping=aes(x=Height,y=Type))
```

```{r}
# Determing if the outliars are caused by age or size

#Setting up read.csv
regular_ol_babalone <- read_csv("data/abalone.csv")
regular_ol_babalone
abalone2 <- select(regular_ol_abalone,- X1)

#without jitter
ggplot(data=abalone2)+
  geom_point(mapping=aes(x=Height,y=Rings))

#with jitter
ggplot(data=abalone2)+
  geom_point(mapping=aes(x=Height,y=Rings),position="jitter")
```
According to the plotted data the outliars appear ro be due to large size rather than age as the value of height for the outliars is relativly but the Rings value is low

```{r}
regular_ol_babalone <- read_csv("data/abalone.csv")
regular_ol_babalone
abalone2 <- select(regular_ol_abalone,- X1)

fabalone <- filter(abalone2,Height<0.3)
fabalone

#I have no clue how to do the factoring part. Always returns as data unusable by ggplot

#A quick look at the data after filtering
ggplot(data=fabalone)+
  geom_point(mapping=aes(x=Height,y=Rings),position="jitter")

#Just for fun
ggplot(data=fabalone)+
  geom_point(mapping=aes(x=Height,y=Rings,color=Type),position="jitter")


```

## Darters
```{r}
dtr <- read_fwf("data/darter.txt")
dtr
```

