---
title: "HW 08: More Data Wrangling"
author: "Cody Grainger"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Chapter 5: Data transformation

Read [R4ds Chapter 5: Data transformation](https://r4ds.had.co.nz/transform.html), sections 5-7.

**You do not answer any of the questions this week.** Just read and run all of the examples as noted below. Pay attention to the code you enter for the examples. You will be applying the information in future assignments.

# 5.1.1

Load load packages/data

```{r}
library(nycflights13)
library(tidyverse)
```


### Section 5.5: Add new variables with `mutate()`

Type and run all of the examples.

Pay careful attention to the useful creation functions in section 5.5.1. 

#

Adding new columns with the mutate function. Mutate() adds columns to the end of the data set

```{r}
# --
# Creating new columns
flights_sml <- select(flights, 
  year:day, 
  ends_with("delay"), 
  distance, 
  air_time
)
mutate(flights_sml,
  gain = dep_delay - arr_delay,
  speed = distance / air_time * 60
)

# --
# Referring to new columns created
mutate(flights_sml,
  gain = dep_delay - arr_delay,
  hours = air_time / 60,
  gain_per_hour = gain / hours
)

# --
#  Transmute()
#When you oly want to keep the new variable
transmute(flights,
  gain = dep_delay - arr_delay,
  hours = air_time / 60,
  gain_per_hour = gain / hours
)
```
#

# 5.5.1

Useful creation Functions

Functions for mutate() must be vectorized aka it will only input and output vector values

Arithmetic operators - +, -, *, /, ^
   Ex: "air_time / 60, hours * 60 + minute"
   
Arithmetic functions with aggregate functions
   Ex: x/sum(x) for proportion of a total, y-mean(y) computes difference from the mean
   
Modular arithmetic %/% (integer division) and %% (remainder), where x==y*(x%/%y)+(x%%y)

```{r}
#Modular arithmetic
transmute(flights,
  dep_time,
  hour = dep_time %/% 100,
  minute = dep_time %% 100
)
```
Logs can be used - log(), log2(), log10()

Offsets allow leading or lagging values. lead(), lag()
   Ex: x-lag(x) for running differences
   . x!=lag(x) to look for values changed. useful with group_by()
```{r}
#lag and lead
x <- c(1:10)
x
lag(x)
lead(x)
```

cumulative and rolling aggregates
Sumns, products, mins, maxes, means(dplyr only)
cumsum(), cumprod(), cummin(), cummax(), cummean()
```{r}
x <- c(1:10)
x
cumsum(x)
cummean(x)
```

Ranking
min_rank() gives smallest values the smallest rank. desc(x) to give the largest values first
```{r}
#Ranking
y <- c(1, 2, 2, NA, 3, 4)
y
min_rank(y)
min_rank(desc(y))
```
Ranking variants
   row_number(), dense_rank(), percent_rank(), cume_dist(), ntile()
```{r}
# Ranking variants
y <- c(1, 2, 2, NA, 3, 4)
y

row_number(y)
dense_rank(y)
percent_rank(y)
cume_dist(y)
```


### Section 5.6: Grouped summaries with `summarise()`

Read and run all of the examples. Pay careful attention to the useful summary functions. Unless you really like baseball, skip the baseball part of 5.6.3.

#

# 5.6

summarise() collapses a data frame into a single row

group_by() analyzes individual groups of the dataset
```{r}
#Summarize without using group_by()
summarise(flights, delay = mean(dep_delay, na.rm = TRUE))

#Using group_by()
by_day <- group_by(flights, year, month, day)
summarise(by_day, delay = mean(dep_delay, na.rm = TRUE))

```

# 5.6.1
Combining multipl eoperations with the pipe
```{r}
#Example without pipes
by_dest <- group_by(flights, dest)
delay <- summarise(by_dest,
  count = n(),
  dist = mean(distance, na.rm = TRUE),
  delay = mean(arr_delay, na.rm = TRUE)
)
delay <- filter(delay, count > 20, dest != "HNL")

delay
#with corresponding plot
ggplot(data = delay, mapping = aes(x = dist, y = delay)) +
  geom_point(aes(size = count), alpha = 1/3) +
  geom_smooth(se = FALSE)
```

In prparing the data,
  Group flights by destination
  summarize to compute distance, avg delay, and #of flights
  
  originally multiple data frames need a name but not with pipes!
```{r}
#Using pipes
delays <- flights %>% 
  group_by(dest) %>% 
  summarise(
    count = n(),
    dist = mean(distance, na.rm = TRUE),
    delay = mean(arr_delay, na.rm = TRUE)
  ) %>% 
  filter(count > 20, dest != "HNL")

delays
```
Pipe notes: x %>% f(y) = f(x, y), and x %>% f(y) %>% g(z) = g(f(x, y), z)

# 5.6.2
Missing values

na.rm
(Missing inputs produce missing outputs)

if we dont set it...
```{r}
#no na.rm
flights %>% 
  group_by(year, month, day) %>% 
  summarise(mean = mean(dep_delay))

#na.rm
flights %>% 
  group_by(year, month, day) %>% 
  summarise(mean = mean(dep_delay, na.rm = TRUE))
```
```{r}
# The missing values represent cancelled flights. Removing cancelled flights 

not_cancelled <- flights %>% 
  filter(!is.na(dep_delay), !is.na(arr_delay))

not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(mean = mean(dep_delay))
```

# 5.6.4
Useful summary functions

subsetting
```{r}
not_cancelled <- flights %>% 
  filter(!is.na(dep_delay), !is.na(arr_delay))

#subsetting
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(
    avg_delay1 = mean(arr_delay),
    avg_delay2 = mean(arr_delay[arr_delay > 0]) # the average positive delay
  )
```

Measure of spread 
sd(x) - standard deviation, IQR(x) - interquartile range, mad(x) - median absolute deviation
```{r}
#Spread
not_cancelled %>% 
  group_by(dest) %>% 
  summarise(distance_sd = sd(distance)) %>% 
  arrange(desc(distance_sd))
```

Mrasures of rank
min(x), quantile(x,0.25) - find values of x greater than 25% of the values but less than the remaining 75%, max(x)
```{r}
#rank
#When do the first and last flights leave?
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(
    first = min(dep_time),
    last = max(dep_time)
  )
```

Measure position
first(x), nth(x,2), last(x)
```{r}
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(
    first_dep = first(dep_time), 
    last_dep = last(dep_time)
  )

#Filtering gives all of the original variables
not_cancelled %>% 
  group_by(year, month, day) %>% 
  mutate(r = min_rank(desc(dep_time))) %>% 
  filter(r %in% range(r))
```

Counts
n() for returning size of current group
sum(!is.na(x)) to count non-missing values
n_distinct(x) counts number of distinct values
```{r}
#Counts
not_cancelled %>% 
  group_by(dest) %>% 
  summarise(carriers = n_distinct(carrier)) %>% 
  arrange(desc(carriers))

#Simple count
not_cancelled %>% 
  count(dest)

#Optionally providing the weight variable
not_cancelled %>% 
  count(tailnum, wt = distance)
```

Counts and proportions of logical values
When used with numeric functions. TRUE is converted to 1 and FALSE to 0
```{r}
#Number of TRUE values for n_early each day
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(n_early = sum(dep_time < 500))

#Proportion of flights delayed more than an hour
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(hour_prop = mean(arr_delay > 60))
```

#5.6.5
Grouping by multiple variables
each summary peels off one level of grouping
```{r}
#Sumamrizes multiple levels of data

daily <- group_by(flights, year, month, day)
per_day   <- summarise(daily, flights = n())
per_day

per_month <- summarise(per_day, flights = sum(flights))
per_month

per_year  <- summarise(per_month, flights = sum(flights))
per_year

```

#5.6.6 
Ungrouping
ungroup() ungroups previously created groups
```{r}
# No longer grouped by date
daily %>%
  ungroup() %>%
  summarise(flights = n())

```


### Section 5.7: Grouped mutates (and filters)

Read and run all of the examples.

Using group_by with operators mutate() and filter()

```{r}
#Filter to find the worst members of each group
flights_sml %>% 
  group_by(year, month, day) %>%
  filter(rank(desc(arr_delay)) < 10)

#Find all groups bigger than a threshold
popular_dests <- flights %>% 
  group_by(dest) %>% 
  filter(n() > 365)
popular_dests

#Per group metrics
popular_dests %>% 
  filter(arr_delay > 0) %>% 
  mutate(prop_delay = arr_delay / sum(arr_delay)) %>% 
  select(year:day, dest, arr_delay, prop_delay)
```




