---
title: "HW_10"
author: "Cody Grainger"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

### Hw_10: Data Visualization 2

##Import and wrangle

Loading initial Libraries
```{r}
library(here)
library(tidyverse)
library(GGally)
library(patchwork)
```
Importing the 'sparrows.txt' data and assigning it to a variable
```{r}
sparrows_raw <- read_tsv(here("data","sparrows.txt"))

sparrows_raw
```
Creating two new columns for the sparrows_raw data and creating a new tibble. setting the row order will create a variable to label the rows 1:1265 and the "." references the original dataframe for reference in the pipe
```{r}
sparrows <- sparrows_raw %>%
  mutate(box_group="sparrows",
         row_order=1:nrow(.))

view(sparrows)
```
The rest of the notebook will reference Zuur et al.

## Figure 2: Outliars

A boxplot for the sparrows data to observe possible outliars
```{r}
sparrows %>%
  ggplot(aes(x=box_group,
             y=wingcrd))+
  geom_boxplot(width=0.25)+
  labs(x=NULL,
       y="Wing length (mm)")
```

