---
title: "HW_10"
author: "Cody Grainger"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

### Hw_10: Data Visualization 2

##Import and wrangle

Loading initial Libraries
```{r}
library(here)
library(tidyverse)
library(GGally)
library(patchwork)
```
Importing the 'sparrows.txt' data and assigning it to a variable
```{r}
sparrows_raw <- read_tsv(here("data","sparrows.txt"))

sparrows_raw
```
Creating two new columns for the sparrows_raw data and creating a new tibble. setting the row order will create a variable to label the rows 1:1265 and the "." references the original dataframe for reference in the pipe
```{r}
sparrows <- sparrows_raw %>%
  mutate(box_group="sparrows",
         row_order=1:nrow(.))

view(sparrows)
```
The rest of the notebook will reference Zuur et al.

## Figure 2: Outliars

A boxplot for the sparrows data to observe possible outliars
```{r}
plot_box <- sparrows %>%
  ggplot(aes(x=box_group,
             y=wingcrd))+
  geom_boxplot(width=0.25)+
  labs(x=NULL,
       y="Wing length (mm)")

plot_box
```

For better assessment of the full range of values a cleveland plot can be used
```{r}
plot_cleveland <- sparrows %>%
  ggplot(aes(x= wingcrd,y=row_order))+
  geom_point()+
  labs(x="Wing length (mm)",
       y="Order of the data")

plot_cleveland
```
using the patchwork package to put the plots together side by side
```{r}
plot_box+plot_cleveland
```
Flipping the cleveland plot for better comparison with the boxplot
```{r}
plot_box+plot_cleveland+coord_flip()
```

## Figure 3: Outliars

Creating 6 scatter plots and combining them with the patchwork package. ncol set to 3
```{r}
#culmen
p_culmen <- sparrows %>% 
  ggplot(aes(y=row_order))+
  geom_point(aes(x=culmen))+
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

#nalopsi
p_nalospi <- sparrows %>% 
  ggplot(aes(y=row_order))+
  geom_point(aes(x=nalospi))+
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

#wt
p_wt <- sparrows %>% 
  ggplot(aes(y=row_order))+
  geom_point(aes(x=wt))+
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

#wingcrd
p_wingcrd <- sparrows %>% 
  ggplot(aes(y=row_order))+
  geom_point(aes(x=wingcrd))+
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

#tarsus
p_tarsus <- sparrows %>% 
  ggplot(aes(y=row_order))+
  geom_point(aes(x=tarsus))+
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

#head
p_head <- sparrows %>% 
  ggplot(aes(y=row_order))+
  geom_point(aes(x=head))+
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())


plot2 <- p_culmen+p_nalospi+p_wt+p_wingcrd+p_tarsus+p_head+plot_layout(ncol=3)

plot2
```

## Figure 5: distribution

assigning individuals to a bin. bin width set to 0.5 for the sparrows data. then using facet_wrap to alphabetically arrange the plots by month for plot 2. Additionally a density plot is created and colored according to months jun, july,august
```{r}
#Histogram
h1 <- sparrows %>%
  filter(Month %in% 6:8) %>% 
  ggplot()+
  geom_histogram(aes(x=wt),
                 binwidth=0.5,
                 boundary=0,
                 closed="right",
                 color="black")+
  scale_x_continuous(breaks=seq(14,28,by=2))
h1

#Histogram2: Faceted Histogram
h2 <- sparrows %>%
  filter(Month %in% 6:8) %>% 
  mutate(Month=case_when(
    Month==6~"June",
    Month==7~"July",
    Month==8~"August")) %>% 
  ggplot()+
  geom_histogram(aes(x=wt),
                 binwidth=0.5,
                 boundary=0,
                 color="black")+
  scale_x_continuous(breaks=seq(14,28,by=2))+
  facet_wrap(~Month,ncol=1)
h2

#density plot
h3 <- sparrows %>%
  filter(Month %in% 6:8) %>% 
  mutate(Month=case_when(
    Month==6~"June",
    Month==7~"July",
    Month==8~"August")) %>% 
  ggplot()+
  geom_density(aes(x=wt,fill=Month,alpha=0.30),
                 binwidth=0.5,
                 boundary=0,
                 color="black")+
  scale_x_continuous(breaks=seq(14,28,by=2))

h3

```
Adding the 2 plots together
```{r}
h1+h2
```
using ggscatmat many scatter plots can be created quickly
```{r}
sparrows %>% ggscatmat(columns=1:7)
```

## Figure 11: interactions

The code below shows the interaction between sex and season. only months 5-9 are used and immature individuals are removed.mutate assigns names to the months according to the months number in the data set.a trend line is added with the se shader removed.
```{r}
month_order=c("May","June","July","August","September")
sparrows %>% 
  filter(Month %in% 5:9,
         Sex !=0) %>% 
  mutate(Month=case_when(
    Month==5 ~ "May",
    Month==6 ~ "June",
    Month==7 ~ "July",
    Month==8 ~ "August",
    Month==9 ~ "September"),
    Month=factor(Month,levels=month_order,ordered=TRUE),
    Sex=ifelse(Sex==4,"Male","Female")) %>% 
  ggplot(aes(x=wingcrd,
             y=wt))+
  geom_point()+
  geom_smooth(method="lm",se=FALSE)+
  facet_grid(Sex~Month)
```

