---
title: "HW_13"
author: "Cody Grainger"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

### Libraries

Initial libraries
```{r}
library(here)
library(raster)
library(sf)
library(tidyverse)
library(ggthemes)
library(patchwork)
library(maptools)
library(ggmap)
library(marmap)

```

#### Part 1: Outline maps

### Notes

The end product for this section is a U.S. map with missouri filled in black and 3 other states filled in gray.

packages 'map_data' and 'subset' are used to access specific regions (states)

```{r}
# Outline of the 48 contiguous states
usa <- map_data(map="state")

# Outline of just Missouri
us_missouri <- map_data(map="state",
                        region="missouri")

# Outline of Alabama, Florida, Mississippe, and North Carolina
us_other <- subset(x=usa,
                   subset=region %in% c("alabama",
                                        "florida",
                                        "mississippi",
                                        "north carolina"))
```

observing the structure of the usa data frame
```{r}
str(usa)
head(usa)
```

Plotting the 'us_other' states. the fill is set to "gray"
```{r}
ggplot()+
  geom_polygon(data=us_other,
               aes(x=long,y=lat,
                   group=group),
               fill="gray")
```

The code below will add missouri with the fill set to "black" to the original plot.
```{r}
ggplot()+
  geom_polygon(data=us_other,
               aes(x=long,y=lat,
                   group=group),
               fill="gray") +
  geom_polygon(data=us_missouri,
               aes(x=long,y=lat,
                   group=group),
               fill="black")
```

The outline of the US mainland is added
```{r}
ggplot()+
  geom_polygon(data=us_other,
               aes(x=long,y=lat,
                   group=group),
               fill="gray") +
  geom_polygon(data=us_missouri,
               aes(x=long,y=lat,
                   group=group),
               fill="black") +
  geom_polygon(data=usa,
               aes(x=long,y=lat,
                   group=group),
               fill=NA,
               color="black")
```

# US map
 
Now adding the map theme from 'ggthemes' and a coord_fixed layer set to 1.3
```{r}
us_map <- ggplot()+
  geom_polygon(data=us_other,
               aes(x=long,y=lat,
                   group=group),
               fill="gray") +
  geom_polygon(data=us_missouri,
               aes(x=long,y=lat,
                   group=group),
               fill="black") +
  geom_polygon(data=usa,
               aes(x=long,y=lat,
                   group=group),
               fill=NA,
               color="black") +
  theme_map() +
  coord_fixed(1.3)

us_map
```

# Missouri map

A map of missouri divided by counties with scott county filled in black
```{r}
missouri <- map_data("county","missouri")
mo_scott <- subset(missouri, subregion %in% c("scott"))

mo_map <- ggplot() +
  geom_polygon(data=missouri,aes(x=long,y=lat,group=group),fill=NA,color="black") +
  geom_polygon(data=mo_scott,aes(x=long,y=lat,group=group),fill="black") +
  theme_map() +
  coord_fixed(1.3)

mo_map
```

using the patchwork package to set the us map and missouri map side by side. width of the us map is set to 1.5 times wider
```{r}
us_map + mo_map + plot_layout(ncol=2,widths=c(1.5,1))
```

##Outline maps: your turn

California is filled in black on the us_map. San Diego county will be filled in black and Glenn and Contra Costa counties will be shaded gray. the maps are put side by side with patchwork. patchwork width set to 1.8
```{r}
# Outline of the 48 contiguous states
usa <- map_data(map="state")

# Outline of just California
us_california <- map_data(map="state",
                        region="california")


us_map <- ggplot()+
  geom_polygon(data=us_california,
               aes(x=long,y=lat,
                   group=group),
               fill="black") +
  geom_polygon(data=usa,
               aes(x=long,y=lat,
                   group=group),
               fill=NA,
               color="black") +
  theme_map() +
  coord_fixed(1.3)

us_map

california <- map_data("county","california")
ca_san_diego <- subset(california, subregion %in% c("san diego"))
ca_other <- subset(california, subregion %in% c("glenn","contra costa"))
  
ca_map <- ggplot() +
  geom_polygon(data=california,aes(x=long,y=lat,group=group),fill=NA,color="black") +
  geom_polygon(data=ca_san_diego,aes(x=long,y=lat,group=group),fill="black") +
  geom_polygon(data=ca_other,aes(x=long,y=lat,group=group),fill="grey") +
  theme_map() +
  coord_fixed(1.3)

ca_map

us_map + ca_map + plot_layout(ncol=2,widths=c(1.8,1))
```

#

#### Part 2: Choropleth maps

