---
title: "HW_13"
author: "Cody Grainger"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

### Libraries
```{r}
library(tidyverse)
library(here)
library(ggthemes)
library(gghighlight)
```

#### Part 1: DNA of Strings

Importing "beta_gliadin.txt" with read_lines() and skipping the first line. str_replace_all() is used to remove the white spaces and numbers and the result is assigned to "beta". A new variable is created to store the collapsed version of "beta"
```{r}
#Importing the raw data
beta_raw <- read_lines(here("data","beta_gliadin.txt"),skip=1)

#Removing blank spaces
beta <- str_replace_all(beta_raw, "\\s","")

#Removing numbers
(beta <- str_replace_all(beta,"\\d",""))

#using the collapse argument of str_c()
(beta_c <- str_c(beta,collapse=""))
```

Setting the letters to uppercase and finding the string length
```{r}
#Setting the leters to capital

(beta_c <- str_to_upper(beta_c))

str_length(beta_c)

```

Detecting the number of start codon sequences with str_count(). Stop codons TGA and TAA are also displayed.
```{r}
#Count for ATG
#Start codons
str_count(beta_c,"ATG")

#Count for TGA and TAA
#Stop codons
str_count(beta_c,"TGA|TAA")
```

Creating variables CAAT_box and TATA_box and searching for them in the sequence
```{r}
CAAT_box <- "CCAAT"
TATA_box <- "CTATTA"

str_count(beta_c,CAAT_box)

str_count(beta_c,TATA_box)
```

Creating regex search strings to seach for specific sequences that begin and end with "CCAAT","ATG" and "CTATAA","ATG". The found sequences are extracted and printed. Finding seq_length() as well.
```{r}
"CAAT"
#CAAT
#Assigning search string
caat_seq <- "CCAAT.{100,150}ATG"
#count for number of times this defined start and stop of sequence shows up
str_count(beta_c,caat_seq)
#Extracting the sequence
(caat_seqs_all <- str_extract_all(beta_c, caat_seq,simplify=TRUE))
#Length of the extracted sequence
str_length(caat_seqs_all)


#Divider
"---------------------"

"TATA"

#TATA
#Assigning search string
tata_seq <- "CTATAA.{50,100}ATG"
#count for number of times this defined start and stop of sequence shows up
str_count(beta_c,tata_seq)
#Extracting the sequence
(tata_seq_all <- str_extract_all(beta_c, tata_seq,simplify=TRUE))
#Length of the extracted sequence
str_length(tata_seq_all)
```
Using str_which() to find which of the 2 CAAT sequences contains the TATA sequence. Assigning the correct result to a new variable.
```{r}
#Finding which caat sequence contains tata_seq
str_which(caat_seqs_all,tata_seq)

#Extracting the above sequence into a new variable
(promotor <- str_extract(beta_c,caat_seq))
str_length(promotor)
```


#### Part 2: Ants

importing the ants.csv data
```{r}
(ants_raw <- read_csv(here("data","ants.csv")))
```

Tidying the data using regex instead of hard coding all of the names.
```{r}
ants <- ants_raw %>%
  pivot_longer(names_to = "species", values_to = "abundance", matches("[A-Z].[a-z]", ignore.case = FALSE))

ants
```

Using mutate to change the species name into one that uses the first 4 digits of the genus and the first 4 of species.
```{r}
ants2 <- ants %>%
  mutate(species = str_replace(species, "[a-z].[a-z]", "[^A-Z]{1)+[a-z]{3}+"))

ants2
```

#### Part 3

Importing the data "flower_size.csv" and wrangling the data to remove na. columns "Species", "Corolla length (cm)", "Corolla width throat (cm)" are selected for and summarise is used to create new columns.
```{r}
flowers_raw <- read_csv(here("data","flower_size.csv"))

flowers_raw %>%
  drop_na("Flower number")

flowers <- select(flowers_raw, "Species", "Corolla length (cm)", "Corolla width throat (cm)") %>%
  rename("cor_length"="Corolla length (cm)",
         "cor_width"="Corolla width throat (cm)") %>%
  drop_na()

(flowers <- flowers %>%
  separate(Species,
           into=c("genus","species"),
           sep=" "))

flowers %>%
  group_by(genus,species) %>%
  summarise(mean_length=mean(cor_length),
            mean_width=mean(cor_width),
            max_length=max(cor_length),
            max_width=max(cor_width))
```

Adding the second data set for this part.
```{r}
pollinators_raw <- read_csv(here("data","flower_pollinators.csv"))

pollinators <- pollinators_raw %>%
  select(Species,Color,Pollinator) %>%
  rename("color"="Color",
         "pollinator"="Pollinator") %>%

  drop_na()

(pollinators <- pollinators %>%
  separate(Species,
           into=c("genus","species"),
           sep=" "))

(pollinators <- pollinators %>%
  separate(pollinator,
           into="syndrome",
           sep=",| |;",
           extra="drop"))
```

Joining the two data sets
```{r}
flowers %>%
  inner_join(pollinators, by="species")
```


#### Part 4
