---
title: "HW_13"
author: "Cody Grainger"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

### Libraries
```{r}
library(tidyverse)
library(here)
library(ggthemes)
library(gghighlight)
```

#### Part 1: DNA of Strings

Importing "beta_gliadin.txt" with read_lines() and skipping the first line. str_replace_all() is used to remove the white spaces and numbers and the result is assigned to "beta". A new variable is created to store the collapsed version of "beta"
```{r}
#Importing the raw data
beta_raw <- read_lines(here("data","beta_gliadin.txt"),skip=1)

#Removing blank spaces
beta <- str_replace_all(beta_raw, "\\s","")

#Removing numbers
(beta <- str_replace_all(beta,"\\d",""))

#using the collapse argument of str_c()
(beta_c <- str_c(beta,collapse=""))
```

Setting the letters to uppercase and finding the string length
```{r}
#Setting the leters to capital

(beta_c <- str_to_upper(beta_c))

str_length(beta_c)

```

Detecting the number of start codon sequences with str_count(). Stop codons TGA and TAA are also displayed.
```{r}
#Count for ATG
#Start codons
str_count(beta_c,"ATG")

#Count for TGA and TAA
#Stop codons
str_count(beta_c,"TGA|TAA")
```

